---
title: "Weight Lifting Quality Assessment"
author: "Shiraz"
date: "03 April 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary
One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. The goal of this report is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they performed the exercise (i.e. the quality of execution). 

This report utilises machine learning algorithms (specifically random forests, generalized boosted regression and linear discriminant analysis) to predict the class of an activity (i.e. whether it is performed correctly or involved a common error).

## Data Processing

Load the data into separate training and testing datasets

```{r}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

---

Reference: 
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz44lxKNFDx